---
title: "p8105_hw2_ml4418"
author: "Mengyuan Li"
date: "9/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
```

# problem 1-part 1
```{r}
MrTrash_data = read_excel("trash_data.xlsx", sheet = 1, range = cell_cols("A:N"))
MrTrash_data = janitor::clean_names(MrTrash_data) %>%
  drop_na()

mutate(MrTrash_data, sports_balls = as.integer(round(sports_balls,0)))
```


# problem 1-part 2
```{r}
pre2018_data = read_excel("trash_data.xlsx", sheet = 3, skip = 1)
pre2018_data = janitor::clean_names(pre2018_data) %>%
  drop_na()
pre2018_data = mutate(pre2018_data,year = "2018")
  
pre2017_data = read_excel("trash_data.xlsx", sheet = 4, skip = 1)
pre2017_data = janitor::clean_names(pre2017_data) %>%
  drop_na()
pre2017_data = mutate(pre2017_data, year = "2017")

pre_data = bind_rows(pre2018_data, pre2017_data) %>% 
  mutate(month = month.name[month])
```

Conclusion: In Mr.Trash Wheel, the total number of observations is `r nrow(MrTrash_data)`. In precipitation dataset, the total number of observations in 2018 is `r nrow(pre2018_data)`, and the total number of observations in 2017 is `r nrow(pre2017_data)`. Specifically, the total precipitation in 2018 is `r sum(pull(pre2018_data, 'total'))`. Also, the median number of sports balls in a dumpster in 2017 is `r median(pull(MrTrash_data, "sports_balls"))`

# problem 2-clean data in pols-month
```{r}
pols_month_data = read.csv("pols-month.csv") %>%
  separate(mon, into = c("month", "day", "year"), "/") %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = month.name[month])
 
prez_gop = filter(pols_month_data, prez_gop == 1 ) %>%
  mutate(president = "gop")
prez_dem = filter(pols_month_data, prez_dem == 1) %>%
  mutate(president = "dem")

pols_month_data = bind_rows(prez_gop, prez_dem) %>%
  select(-prez_gop, -prez_dem , -day)
```

# problem 2-clean date in snp
```{r}
snp_data = read.csv("snp.csv") %>%
  separate(date, into = c("month","day","year"), "/") %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, -day, close)
```

# problem 2-tidy unemployment data
```{r}
un_data = read.csv("unemployment.csv")
tidy_un_data = 
  pivot_longer(un_data,
               Jan:Dec,
               names_to = "month",
               values_to = "unempolyment") 
tidy_un_data = janitor::clean_names(tidy_un_data) %>%
  mutate(year = as.character(year))
  


join_data = 
  left_join(pols_month_data, snp_data, by = c("month","year"))
final_data = 
  left_join(join_data, tidy_un_data, by = c("month","year"))
```
  
Conclusion: The dimension of snp dataset is `r nrow(snp_data)`* `r ncol(snp_data)`. The dimension of pols-month dataset is `r nrow(pols_month_data)` * `r ncol(pols_month_data)`. The dimension of unemployment dataset is `r nrow(tidy_un_data)` * `r ncol(tidy_un_data)`. The dimension of the final dataset is `r nrow(final_data)` * `r ncol(final_data)`. The range of years in snp dataset is from `r min(pull(snp_data, "year"))` to `r max(pull(snp_data, "year"))`. The range of years in pols-month dataset is from `r min(pull(pols_month_data, "year"))` to `r max(pull(pols_month_data, "year"))`. The range of years in unempolyment dataset is from `r min(pull(tidy_un_data, "year"))` to `r max(pull(tidy_un_data, "year"))`. THe rangeof years in the final dataset is from `r min(pull(final_data, "year"))` to `r max(pull(final_data, "year"))`.


# problem 3-part 1
```{r}
popular_baby_names = read.csv("Popular_Baby_Names.csv")
popular_baby_names = janitor::clean_names(popular_baby_names)
popular_baby_names = unique(popular_baby_names)
```
  

# problem 3-part 2
```{r}
olivia_name_data = filter(popular_baby_names, child_s_first_name == "Olivia") %>%
  select(ethnicity, year_of_birth, rank)
olivia_name_data = pivot_wider(olivia_name_data, 
                               ethnicity:rank, 
                               names_from = "year_of_birth",
                               values_from = "rank")
```


# problem 3-part 3
```{r}
popular_male_names = filter(popular_baby_names, gender == "MALE") %>%
  select(year_of_birth, child_s_first_name, count)

popular_male_names_2016 = filter(popular_male_names, year_of_birth == "2016") 
popular_male_names_2016 = filter(popular_male_names_2016, count == max(count))

popular_male_names_2015 = filter(popular_male_names, year_of_birth == "2015")
popular_male_names_2015 = filter(popular_male_names_2015, count == max(count))

popular_male_names_2014 = filter(popular_male_names, year_of_birth == "2014")
popular_male_names_2014 = filter(popular_male_names_2014, count == max(count))

popular_male_names_2013 = filter(popular_male_names, year_of_birth == "2013")
popular_male_names_2013 = filter(popular_male_names_2013, count == max(count))

popular_male_names_2012 = filter(popular_male_names, year_of_birth == "2012")
popular_male_names_2012 = filter(popular_male_names_2012, count == max(count))

popular_male_names_2011 = filter(popular_male_names, year_of_birth == "2011")
popular_male_names_2011 = filter(popular_male_names_2011, count == max(count))

most_popular_male_name = bind_rows(popular_male_names_2016,
                                   popular_male_names_2015,
                                   popular_male_names_2014,
                                   popular_male_names_2013,
                                   popular_male_names_2012,
                                   popular_male_names_2011)
```


# problem 3- part 4
```{r}
white_males_2016 = filter(popular_baby_names, 
                          gender == "MALE",
                          ethnicity == "WHITE NON HISPANIC",
                          year_of_birth == "2016")
white_males_2016_plot = ggplot(white_males_2016,
                               aes(x = rank, y = count)) + geom_point(aes(color = "lightpink"))
```

  
  
